Equalizer\\Programming and User Guide

### 2010-11-03 Developer/ProgrammingGuide/paper.tex
Error codes: Programming Guide update
[ ] May break build
[ ] May break existing applications (see CHANGES.txt)
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications using
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.8 for Equalizer 0.9.1+svn\\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    1.7.3   & Feb 02, 2010 & Revision for Equalizer 0.9.1\\
    1.7.4   & Feb 22, 2010 & Add \sref{sSlaveCommit}\\
    1.7.5   & Jun 03, 2010 & Add \sref{sClusterSetup}\\
    1.7.6   & Aug 23, 2010 & General update for fabric changes\\
    1.7.7   & Oct 01, 2010 & SystemPipe and SystemWindow renaming\\
    1.7.8   & Oct 04, 2010 & Document GPU affinity workaround hint\\
    \hline
  \end{tabularx}}
\clearpage

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/processes.pdf}
  {\caption{\label{fProcesses}Equalizer Processes}}
\end{wrapfigure}
The render client implements the rendering part of an application. Its
execution is passive, it has no main loop and is completely driven by
Equalizer. It executes the rendering tasks received from the server by
calling the appropriate task methods (see \sref{sTaskMethods}) in the
correct thread and context. The application either implements the task
methods with application-specific code or uses the default methods
provided by Equalizer.

The application can also be a rendering client, in which case it can
also contribute to the rendering. If it does not implement any render
client code, it is reduced to be the application's `master' process
without any OpenGL windows and 3D rendering.

@@

Stereo compounds, as shown in \fig{fStereoCmp}\footnote{3D model
  courtesy of Stereolithography Archive at Clemson University.}, assign
each eye pass to individual rendering units. The resulting images are
copied to the appropriate stereo buffer. This mode supports a variety of
stereo modes, including active (quad-buffered) stereo, anaglyphic stereo
and auto-stereo displays with multiple eye passes.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\label{fStereoCmp}\small A Stereo Compound}}
\end{wrapfigure}
Eye compounds in Equalizer are configured using the \textsf{eye}
parameter, limiting the child to render the \textsf{[ LEFT ]} or
\textsf{[ RIGHT ]} eye. Each child compound uses an output frame, which
is connected to an input frame on the destination channel. The
destination channel can also be used to render an eye pass, in which case
it renders in the correct stereo buffer and no output frame is needed.

@@

Pixel compounds are similar to 2D compounds. The frusta of the
source rendering units are modified so that each unit renders an evenly
distributed subset of pixels, as shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

As 2D compounds, pixels compounds have low, constant IO requirements for
the pixel transfers during recomposition.

@@

    config->setInitData( _initData );
    if( !config->init( ))
    {
        EQERROR << "Config initialization failed: " 
                << config->getErrorMessage() << endl;
        server->releaseConfig( config );
        disconnectServer( server );
        return EXIT_FAILURE;
    }

@@

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/threads.pdf}
  {\caption{\label{fThreads}Threads within one Node Process}}
\end{wrapfigure}
The application or node main thread is the primary thread of each
process and executes the \textsf{main} function. The application and
render clients initialize the local node for communications with other
nodes, including the server, using \textsf{Client::initLocal}.

During this initialization, Equalizer creates and manages two threads
for communication, the receiver thread and the command thread. Normally
no application code is executed from these two threads.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frameSync.pdf}
  {\caption{\label{fFrameSync}Per-Node Frame Synchronization}}
\end{wrapfigure}
The application has extended control over the task synchronization
during a frame. Upon \textsf{Config::startFrame}, Equalizer invokes the
\textsf{frameStart} task methods of the various entities. The entities
unlock all their children by calling \textsf{startFrame}, e.g.,
\textsf{Node::frameStart} has to call \textsf{Node::startFrame}
to unlock the pipe threads. Note that certain \textsf{startFrame} calls,
e.g., \textsf{Window::startFrame}, are currently empty since the
synchronization is implicit due to the sequential execution within the
thread.

@@

\fig{fFrameSync} illustrates the local frame synchronization.

@@

Since window initialization is notoriously error-prone and hard to
debug, the sample implementation propagates the reason for errors
from the render clients back to the application. The \textsf{Pipe} and
\textsf{Window} classes have a \textsf{setErrorMessage} method, which is
used to set an error string. This string is passed to the
\textsf{Config} instance on the application node, where it can be
retrieved using \textsf{getErrorMessage}.

@@

@@

    if( success && !_xDrawable )
    {
        _window->setErrorMessage( 
            "configInitGLXDrawable did not set a X11 drawable");
        return false;
    }

@@

{\footnotesize\begin{lstlisting}
bool WGLWindow::configInit()
{
    if( !initWGLAffinityDC( ))
    {
        _window->setErrorMessage( "Can't create affinity dc" );
        return false;
    }

@@

    if( !_wglDC )
    {
        exitWGLAffinityDC();
        setWGLDC( 0, WGL_DC_NONE );
        _window->setErrorMessage(
            "configInitWGLDrawable did not set a WGL drawable" );
        return false;
    }

@@

    setWGLContext( context );
    makeCurrent();
    _initGlew();

@@

    if( !joinNVSwapBarrier( ))
    {
        _window->setErrorMessage( "Joining NV_swap_group failed" );
        return false;
    }

@@

\begin{wrapfigure}{r}{.612\textwidth}
  \includegraphics[width=.612\textwidth]{images/viewUML.pdf}
  {\caption{\label{fViewUML}UML Hierarchy of \textsf{eqPly::View}}}
\end{wrapfigure}
The render clients can access a slave instance of the view using
\textsf{Channel::getView}. When called from one of the frame task
methods, this method will return the view of the current destination
channel for which the task method is executed. Otherwise it returns the
channel's native view, if it has one. Only destination channels of an
active canvas have a native view.

@@

The most common entity to subclass is the \textsf{View}, since the application
often amends with view-specific application data. A view might have a
distributable user data object, which has to inherit from
\textsf{eq::net::Object} to be committed and synchronized with the associated
object.

@@

The frustum of a view overrides the underlying frustum of the
segments. The frustum of a view does not typically correspond to a
physical projection, and can therefore be changed at runtime. The
frustum of a segment does however correspond to a physical projection,
and is therefore immutable.

@@

\begin{wrapfigure}{l}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewUpdate.pdf}
  {\caption{\label{fViewUpdate}Event Flow during a View Update}}
\end{wrapfigure}
When a window is resized, a \textsf{CHANNEL\_RESIZE} event is eventually
generated. If the corresponding channel has a view,
\textsf{Channel::processEvent} sends a \textsf{VIEW\_RESIZE} event to
the application. This event contains the identifier of the view.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\label{fAssembly}Hierarchy of Assembly Classes}}
\end{wrapfigure}
Readback and assemble operations on frames and images are designed to be
asynchronous. They have a start and finish method for both readback and
assemble to allow the initiation and synchronization of the operation.
Currently, only synchronous readback and assembly using
\textsf{glReadPixels} and \textsf{glDrawPixels} is implemented in the
respective start method of the image. Later versions of Equalizer will
implement asynchronous pixel transfers.

The offset of input and output frames characterizes the position of the
frame data relative to the framebuffer, that is, the \textbf{window's}
lower-left corner. For output frames this is the position of the channel
relative to the window.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}\vspace{-2ex}
\end{wrapfigure}
The \textsf{eVolve} example is a scalable volume renderer. It uses 3D
texture-based volume rendering, where the 3D texture is intersected by
view-aligned slices. The slices are rendered back-to-front and blended
to produce the final image, as shown in \fig{fSlices}\footnote{Volume
  Data Set courtesy of: SFB-382 of the German Research Council (DFG)}.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
DB compounds have the advantage of scaling any part of the volume
rendering pipeline: texture and main memory (smaller bricks for each
channel), fill rate (less samples per channel) and IO bandwidth for
time-dependent data (less data per time step and channel). Since the
amount of texture memory needed for each node decreases linearly, they
make it possible to render data sets which are not feasible to
visualize with any other approach.

@@

For orthographic rendering, determining the compositing order of the
input frames is trivial. The screen-space orientation of the volume
bricks determines the order in which they have to be composited. The
bricks in \textsf{eVolve} are created by slicing the volume along one
dimension. Therefore the range of the resulting frame buffer images,
together with the sorting order, is used to arrange the frames during
compositing. \fig{fBlendOrtho} shows this composition for one view.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netSession.pdf}
  {\caption{\label{fNetSession}Communication between two Sessions}}
\end{wrapfigure}
The \textsf{eq::net::Session} gathers provides higher-level
functionality to a set of nodes. The master instance of a session is
registered to a node, which makes this node the session server and
assigns a node-unique identifier to the session. All other nodes map
their instance of the session to their local listening node using the
session's identifier. One given node can participate in any number of
sessions, both as master and slave. This allows the flexible usage of
sessions to group different aspects of an application.

A session provides the facilitate to allocate session-unique
identifiers. These identifiers can be used by the application to
uniquely address data within the cluster. They are also used by the
session to address distributed objects, which are described in the
following section.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/slaveWrites.pdf}
  {\caption{\label{fSlaveCommit}Slave Commit Communication Sequence}}
\end{wrapfigure}
When using \textsf{commit} on slave object instances, multiple modifications on
the same database may happen simultaneously. In order to avoid excessive
synchronization in the network layer, Equalizer a light-weight approach which
requires the application to avoid or handle conflicting changes on the same
object.

@@

The master instance can \textsf{sync} delta data received from slaves. The data
is synchronized in the order it has been received from the slaves. If a specific
order is needed, it is the application's responsibity to serialize the slave
commits accordingly. When using \textsf{VERSION\_NEXT} as the version parameter
to \textsf{sync}, one slave commit is unpacked. This call may block if no data
is pending. When using \textsf{VERSION\_HEAD} all pending versions are
unpacked. This call will never block.

>>>>

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications using
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.9 for Equalizer 0.9.1+svn\\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    1.7.3   & Feb 02, 2010 & Revision for Equalizer 0.9.1\\
    1.7.4   & Feb 22, 2010 & Add \sref{sSlaveCommit}\\
    1.7.5   & Jun 03, 2010 & Add \sref{sClusterSetup}\\
    1.7.6   & Aug 23, 2010 & General update for fabric changes\\
    1.7.7   & Oct 01, 2010 & SystemPipe and SystemWindow renaming\\
    1.7.8   & Oct 04, 2010 & Document GPU affinity workaround hint\\
    1.7.9   & Nov 03, 2010 & Add \sref{sErrorHandling}\\
    \hline
  \end{tabularx}}
\clearpage

@@

The render client implements the rendering part of an application. Its
execution is passive, it has no main loop and is completely driven by
Equalizer. It executes the rendering tasks received from the server by
calling the appropriate task methods (see \sref{sTaskMethods}) in the
correct thread and context. The application either implements the task
methods with application-specific code or uses the default methods
provided by Equalizer.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/processes.pdf}
  {\caption{\label{fProcesses}Equalizer Processes}}
\end{wrapfigure}
The application can also be a rendering client, in which case it can
also contribute to the rendering. If it does not implement any render
client code, it is reduced to be the application's `master' process
without any OpenGL windows and 3D rendering.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\label{fStereoCmp}\small A Stereo Compound}}
\end{wrapfigure}
Stereo compounds, as shown in \fig{fStereoCmp}\footnote{3D model
  courtesy of Stereolithography Archive at Clemson University.}, assign
each eye pass to individual rendering units. The resulting images are
copied to the appropriate stereo buffer. This mode supports a variety of
stereo modes, including active (quad-buffered) stereo, anaglyphic stereo
and auto-stereo displays with multiple eye passes.

@@

Eye compounds in Equalizer are configured using the \textsf{eye}
parameter, limiting the child to render the \textsf{[ LEFT ]} or
\textsf{[ RIGHT ]} eye. Each child compound uses an output frame, which
is connected to an input frame on the destination channel. The
destination channel can also be used to render an eye pass, in which case
it renders in the correct stereo buffer and no output frame is needed.

@@

Pixel compounds are similar to 2D compounds. While 2D compounds decompose the
screen space into contiguous regions, pixel compounds assign one pixel of a
regular kernel to each resource. The frusta of the source rendering units are
modified so that each unit renders an evenly distributed subset of pixels, as
shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

As 2D compounds, pixels compounds have low, constant IO requirements for
the pixel transfers during recomposition. They are naturally load-balanced for
fill-limited operations, while they do not decompose the geometry processing at
all.

@@

    config->setInitData( _initData );
    if( !config->init( ))
    {
        server->releaseConfig( config );
        disconnectServer( server );
        return EXIT_FAILURE;
    }

@@

\subsection{\label{sErrorHandling}Error Handling}

All Equalizer entities use an error code to report error conditions during
various operations. This error code is reported back from the render processes
to the application config instance. The last error is propagated from the failed
resource to the config object, where it can be queried by the application using
\textsf{Config::getError}.

Each error emitted by Equalizer code has a textual description. This string is
automatically used for printing an \textsf{Error} on an \textsf{std::ostream},
and can be queried from the \textsf{base::ErrorRegistry}, which is accessible
from \textsf{base::Global}.

Applications can register additional error codes and strings. The
\textsf{eVolve} example uses this to report if required OpenGL extensions are
missing. The registration of new error codes is not thread-safe, that is, no
other thread should use the error registry when it is modified. It is therefore
strongly advised to register application-specific errors before
\textsf{eq::init} and clear them after \textsf{eq::exit}:

{\footnotesize\begin{lstlisting}
int main( const int argc, char** argv )
{
    ...
    // 2. Equalizer initialization
    NodeFactory nodeFactory;
    eVolve::initErrors();

    if( !eq::init( argc, argv, &nodeFactory ))
    ...

    eq::exit();
    eVolve::exitErrors();
    return ret;
}
\end{lstlisting}}

When the application returns false from a \textsf{configInit} task method due to
an application-specific error, it should set an error code using
\textsf{setError}. Application-specific errors can have any value equal or
greater than \textsf{eq::ERROR\_CUSTOM}:

{\footnotesize\begin{lstlisting}
/** Defines errors produced by eVolve. */
enum Error
{
    ERROR_EVOLVE_ARB_SHADER_OBJECTS_MISSING = eq::ERROR_CUSTOM,
    ERROR_EVOLVE_EXT_BLEND_FUNC_SEPARATE_MISSING,
    ERROR_EVOLVE_ARB_MULTITEXTURE_MISSING,
    ERROR_EVOLVE_LOADSHADERS_FAILED,
    ERROR_EVOLVE_LOADMODEL_FAILED
};

bool Window::configInitGL( const uint32_t initID )
{
    ...
    if( !GLEW_ARB_shader_objects )
    {
        setError( ERROR_EVOLVE_ARB_SHADER_OBJECTS_MISSING );
        return false;
    }
    ...
}
\end{lstlisting}}

It is not required to register an error string for each application-specific
error. Registering a string will however cause error printing to use this string
instead of using only the numerical error value. Applications may also redefine
existing error strings to their convenience, e.g., for internationalization
purposes:

{\footnotesize\begin{lstlisting}
namespace
{
struct ErrorData
{
    const uint32_t code;
    const std::string text;
};

ErrorData _errors[] = {
    { ERROR_EVOLVE_ARB_SHADER_OBJECTS_MISSING,
      "GL_ARB_shader_objects extension missing" },
    { ERROR_EVOLVE_EXT_BLEND_FUNC_SEPARATE_MISSING,
      "GL_ARB_shader_objects extension missing" },
    { ERROR_EVOLVE_ARB_MULTITEXTURE_MISSING,
      "GL_ARB_shader_objects extension missing" },
    { ERROR_EVOLVE_LOADSHADERS_FAILED, "Can't load shaders" },
    { ERROR_EVOLVE_LOADMODEL_FAILED, "Can't load model" },

    { 0, "" } // last!
};
}

void initErrors()
{
    eq::base::ErrorRegistry& registry = eq::base::Global::getErrorRegistry();

    for( size_t i=0; _errors[i].code != 0; ++i )
        registry.setString( _errors[i].code, _errors[i].text );
}

void exitErrors()
{
    eq::base::ErrorRegistry& registry = eq::base::Global::getErrorRegistry();

    for( size_t i=0; _errors[i].code != 0; ++i )
        registry.eraseString( _errors[i].code );
}
\end{lstlisting}}

@@

The application or node main thread is the primary thread of each
process and executes the \textsf{main} function. The application and
render clients initialize the local node for communications with other
nodes, including the server, using \textsf{Client::initLocal}.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/threads.pdf}
  {\caption{\label{fThreads}Threads within one Node Process}}
\end{wrapfigure}
During this initialization, Equalizer creates and manages two threads
for communication, the receiver thread and the command thread. Normally
no application code is executed from these two threads.

@@

The application has extended control over the task synchronization
during a frame. Upon \textsf{Config::startFrame}, Equalizer invokes the
\textsf{frameStart} task methods of the various entities. The entities
unlock all their children by calling \textsf{startFrame}, e.g.,
\textsf{Node::frameStart} has to call \textsf{Node::startFrame}
to unlock the pipe threads. Note that certain \textsf{startFrame} calls,
e.g., \textsf{Window::startFrame}, are currently empty since the
synchronization is implicit due to the sequential execution within the
thread.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frameSync.pdf}
  {\caption{\label{fFrameSync}Per-Node Frame Synchronization}}
\end{wrapfigure}
\fig{fFrameSync} illustrates the local frame synchronization.

@@

Since window initialization is notoriously error-prone and hard to debug, the
sample implementation propagates the reason for errors from the render clients
back to the application. The \textsf{Pipe} and \textsf{Window} classes have a
\textsf{setError} method, which is used to set an error code. This string is
passed to the \textsf{Config} instance on the application node, where it can be
retrieved using \textsf{getError}. The \sref{sErrorHandling} explains error
handling in more detail.

@@

    // Early glxew init to have function pointers for init code
    glXMakeCurrent( _xDisplay, None, 0 );
    glxewInit();

@@

    if( !success || !_xDrawable )
    {
        if( getError() == ERROR_NONE )
            setError( ERROR_GLXWINDOW_NO_DRAWABLE );
        return false;
    }

    makeCurrent();
    initGLEW();

    if( getIAttribute( Window::IATTR_HINT_DRAWABLE ) == FBO )
        configInitFBO();

@@

{\footnotesize\begin{lstlisting}
bool WGLWindow::configInit()
{
    if( !initWGLAffinityDC( ))
    {
        setError( ERROR_WGL_CREATEAFFINITYDC_FAILED );
        return false;
    }

@@

    if( !_wglDC )
    {
        exitWGLAffinityDC();
        setWGLDC( 0, WGL_DC_NONE );
        setError( ERROR_WGLWINDOW_NO_DRAWABLE );
        return false;
    }

@@

    setWGLContext( context );
    makeCurrent();
    initGLEW();

@@

@@

The render clients can access a slave instance of the view using
\textsf{Channel::getView}. When called from one of the frame task
methods, this method will return the view of the current destination
channel for which the task method is executed. Otherwise it returns the
channel's native view, if it has one. Only destination channels of an
active canvas have a native view.

@@

\begin{wrapfigure}{r}{.612\textwidth}
  \includegraphics[width=.612\textwidth]{images/viewUML.pdf}
  {\caption{\label{fViewUML}UML Hierarchy of \textsf{eqPly::View}}}
\end{wrapfigure}
The most common entity to subclass is the \textsf{View}, since the application
often amends with view-specific application data. A view might have a
distributable user data object, which has to inherit from
\textsf{eq::net::Object} to be committed and synchronized with the associated
object.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewUpdate.pdf}
  {\caption{\label{fViewUpdate}Event Flow during a View Update}}
\end{wrapfigure}
The frustum of a view overrides the underlying frustum of the
segments. The frustum of a view does not typically correspond to a
physical projection, and can therefore be changed at runtime. The
frustum of a segment does however correspond to a physical projection,
and is therefore immutable.

@@

When a window is resized, a \textsf{CHANNEL\_RESIZE} event is eventually
generated. If the corresponding channel has a view,
\textsf{Channel::processEvent} sends a \textsf{VIEW\_RESIZE} event to
the application. This event contains the identifier of the view.

@@

Readback and assemble operations on frames and images are designed to be
asynchronous. They have a start and finish method for both readback and
assemble to allow the initiation and synchronization of the operation.
Currently, only synchronous readback and assembly using
\textsf{glReadPixels} and \textsf{glDrawPixels} is implemented in the
respective start method of the image. Later versions of Equalizer will
implement asynchronous pixel transfers.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\label{fAssembly}Hierarchy of Assembly Classes}}
\end{wrapfigure}
The offset of input and output frames characterizes the position of the
frame data relative to the framebuffer, that is, the \textbf{window's}
lower-left corner. For output frames this is the position of the channel
relative to the window.

@@

The \textsf{eVolve} example is a scalable volume renderer. It uses 3D
texture-based volume rendering, where the 3D texture is intersected by
view-aligned slices. The slices are rendered back-to-front and blended
to produce the final image, as shown in \fig{fSlices}\footnote{Volume
  Data Set courtesy of: SFB-382 of the German Research Council (DFG)}.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
DB compounds have the advantage of scaling any part of the volume
rendering pipeline: texture and main memory (smaller bricks for each
channel), fill rate (less samples per channel) and IO bandwidth for
time-dependent data (less data per time step and channel). Since the
amount of texture memory needed for each node decreases linearly, they
make it possible to render data sets which are not feasible to
visualize with any other approach.

@@

\begin{wrapfigure}{l}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
For orthographic rendering, determining the compositing order of the
input frames is trivial. The screen-space orientation of the volume
bricks determines the order in which they have to be composited. The
bricks in \textsf{eVolve} are created by slicing the volume along one
dimension. Therefore the range of the resulting frame buffer images,
together with the sorting order, is used to arrange the frames during
compositing. \fig{fBlendOrtho} shows this composition for one view.

@@

The \textsf{eq::net::Session} gathers provides higher-level
functionality to a set of nodes. The master instance of a session is
registered to a node, which makes this node the session server and
assigns a node-unique identifier to the session. All other nodes map
their instance of the session to their local listening node using the
session's identifier. One given node can participate in any number of
sessions, both as master and slave. This allows the flexible usage of
sessions to group different aspects of an application.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netSession.pdf}
  {\caption{\label{fNetSession}Communication between two Sessions}}
\end{wrapfigure}
A session provides the facilitate to allocate session-unique
identifiers. These identifiers can be used by the application to
uniquely address data within the cluster. They are also used by the
session to address distributed objects, which are described in the
following section.

@@

When using \textsf{commit} on slave object instances, multiple modifications on
the same database may happen simultaneously. In order to avoid excessive
synchronization in the network layer, Equalizer a light-weight approach which
requires the application to avoid or handle conflicting changes on the same
object.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/slaveWrites.pdf}
  {\caption{\label{fSlaveCommit}Slave Commit Communication Sequence}}
\end{wrapfigure}
The master instance can \textsf{sync} delta data received from slaves. The data
is synchronized in the order it has been received from the slaves. If a specific
order is needed, it is the application's responsibity to serialize the slave
commits accordingly. When using \textsf{VERSION\_NEXT} as the version parameter
to \textsf{sync}, one slave commit is unpacked. This call may block if no data
is pending. When using \textsf{VERSION\_HEAD} all pending versions are
unpacked. This call will never block.
